---
title: "Sparse covariance estimation"
author: "Xuelong Wang"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    keep_tex: true
    fig_width: 12
    fig_height: 3
header-includes:
    - \usepackage{float,amsmath, bbm, siunitx, bm}
    - \usepackage{pdfpages}
    - \floatplacement{figure}{H}
    - \newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = "", message = FALSE, warning = FALSE, echo = FALSE, fig.height = 10)
options(scipen=1, digits=2)

library(MASS)
library(tidyverse)
library(data.table)
library(ggforce)
library(ggpubr)
library(gridExtra)


# heat map function for correlation matrxi

h_map <- function(x,cor = F) {
  if(cor == F){
    Cor <- cor(x) %>% abs(.)
  } else {
    Cor <- x %>% abs(.)
  }
    
  h1 <- ggplot(data = melt(Cor), aes(x=(Var1), y= (Var2), fill=value)) + 
    scale_fill_gradient(low = "white", high = "blue") +
    geom_tile() +
    theme(axis.text.x  = element_text(angle=90))
  h1
}
```

# Motivation
After decorrelation with the information of the historical data, the correlation between the covariate is reduced. However, there are still some correlation coeifficients are large. That may suggest that after using the historical data, the decorrelated data still is not uncorrelated. But the correlation structure becomes a sparse and symmetric. Therefore, we could apply another decorrelation to further reudce the non-zero correlation, so that we may have a better performance on the following variance estimation procedure. 


# Simulation

## Simulation procedure

### Standardization will not change total variance
1. Standardize the X $\tilde{Z}_m = (X-\mu)A_1$
1. Generate the interaction based on the $\tilde{Z}_{int} = \tilde{Z}_m*\tilde{Z}_m$ without the square terms and set $\tilde{Z}_t = (\tilde{Z}_m, \tilde{Z}_{int})$
1. Generate the Y based on the $\tilde{Z}_t$
1. Estimate the $Var(\tilde{Z}_t \beta_t)$ by $Z = \tilde{Z}_tA_2$, where $A_2$ is for decorrelation

Note that the $Var(\tilde{Z}_t \beta_t) = Var(Z\gamma)$,
$A_2 = \hat{\Sigma}^{-1/2}_h$ or $A_2 = \hat{\Sigma}^{-1/2}_h \hat{\Sigma}^{-1/2}_s$


## Decorrelation steps

## two steps decorrelation 
1. Decorrelation by covariance matrix estimated by historical data
1. If after the step 1, the correlation is still large, then we may need a second decorrelation by sparse precision method

## Dpglasso: [The Graphical Lasso: New Insights and Alternatives](https://arxiv.org/pdf/1111.5479.pdf) 

1. An Alternatives for Glasso 
1. Glasso works on $W$ the covariance matrix, but its alg cannot make sure the precision matrix $\Theta$ is positive definite. 
1. Dpglasso can provide both $W$ and $\Theta$ be positive definite


```{r}
source("~/dev/projects/Chen_environmental_study/reports/proposed_GCTA_paper/est_var_analysis/est_combined_data/test_result_sparse_decorr.R")
```



\newpage

## PCB

### None
```{r, PCB None}
result_path <- "~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/decor_method_None_sparse_method_None_result_list_fixed_sub_PCB_structure_un_main_0.5_inter_0.1_n_100_200_500_1000_p_21_rho_e_0.5_decor_FALSE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_total_year_1999_std_PCB_FALSE_c_betam_8_c_betai_2"
file_list_all <- list.files(result_path,full.names = T)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect,
                             structure,
                             decor,
                             x_dist)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n]

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n]
summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

summary_final <- rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE) %>% setorder(., n)
additional
cbind(summary_final)
```


### Hist
```{r, PCB hist}
result_path <- "~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/decor_method_hist_sparse_method_None_result_list_fixed_sub_PCB_structure_un_main_0.5_inter_0.1_n_100_200_500_1000_p_21_rho_e_0.5_decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_total_year_1999_std_PCB_FALSE_c_betam_8_c_betai_2/"
file_list_all <- list.files(result_path,full.names = T)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect,
                             structure,
                             decor,
                             x_dist)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n]

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n]
summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

summary_final <- rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE) %>% setorder(., n)
additional
cbind(summary_final)
```

### Hist + sparse

```{r, PCB hist + sparse}
result_path <- "~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_sparse_decor_report_09_18_2019/decor_method_hist_sparse_method_dgpGLASSO_result_list_fixed_sub_PCB_structure_un_main_0.5_inter_0.1_n_100_200_500_1000_p_21_rho_e_0.5_decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_total_year_1999_std_PCB_FALSE_c_betam_8_c_betai_2/"
file_list_all <- list.files(result_path,full.names = T)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect,
                             structure,
                             decor,
                             x_dist)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n]

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n]
summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

summary_final <- rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE) %>% setorder(., n)
additional
cbind(summary_final)
```

## Chi

### None
```{r, Chi none}
result_path <- "~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/decor_method_None_sparse_method_None_result_list_fixed_sub_chi_structure_un_main_0.5_inter_0.1_n_100_200_500_1000_p_21_rho_e_0.5_decor_FALSE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_total_c_betam_8_c_betai_2_Var_cor"
file_list_all <- list.files(result_path,full.names = T)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect,
                             structure,
                             decor,
                             x_dist)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n]

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n]
summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

summary_final <- rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE) %>% setorder(., n)
additional
cbind(summary_final)

```

### hist
```{r, Chi hist}
result_path <- "~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/decor_method_hist_sparse_method__result_list_fixed_sub_chi_structure_un_main_0.5_inter_0.1_n_100_200_500_1000_p_21_rho_e_0.5_decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_total_c_betam_8_c_betai_2_Var_cor/"
file_list_all <- list.files(result_path,full.names = T)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect,
                             structure,
                             decor,
                             x_dist)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n]

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n]
summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

summary_final <- rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE) %>% setorder(., n)
additional
cbind(summary_final)

```


### hist + sparse
```{r, Chi hist + sparse}
result_path <- "~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/decor_method_hist_sparse_method_Glasso_result_list_fixed_sub_chi_structure_un_main_0.5_inter_0.1_n_100_200_500_1000_p_21_rho_e_0.5_decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_total_c_betam_8_c_betai_2_Var_cor"
file_list_all <- list.files(result_path,full.names = T)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect,
                             structure,
                             decor,
                             x_dist)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n]

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n]
summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

summary_final <- rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE) %>% setorder(., n)
additional
cbind(summary_final)
```

# Mimic the histrocial covariance situation

Based on the previous simulation, we found that if the historical data is from extact the same distribution, then the sparse decorrelation may not be neccessary. That is after doing the second step of decorrelation the variance estimation does not get much improvement. 
So to mimic the situation situation we change the simulation so that there is historical covariance is not perfect. 

## Simulation setup

- $p = 21$
- $X \sim \chi^2_1$
- $Cor(X)$ is the sample covariance of subset of standardized PCB data with $n = 150$  
- $X_h = XB$, where 
$B = \begin{bmatrix}
1 & 1 & 1 & 1 &1\dots \\ 
0 &  1&  1&  1& 1\dots\\ 
0&  & 1 & 1 & 1\dots\\ 
0&  &  &  1& 1\dots\\ 
0&  &  &  & 1\dots
\end{bmatrix}$. 
So that the $X_h$ is a column transfromation of $X$, therefore the its covariance is different from the true value. 

## Decorrelation result

### $X_h = X$
```{r}
## load the dateset 
library(R.utils)
library(MASS)
library(tidyverse)
library(foreach)
library(doRNG)
library(doParallel)
library(gtools) # for rbind based on columns
library(Matrix)
options(warn = 1, error = bettertrace::stacktrace)
setwd("~/dev/projects/Chen_environmental_study/")
sourceDirectory("./R_code/main_fn/",modifiedOnly = FALSE, recursive = TRUE)
sourceDirectory("./R_code/main_fn/method/",modifiedOnly = FALSE, recursive = TRUE)
source("./R_code/simulation_proposed_GCTA/local_helpers.R")
X_orignal <- read.csv("~/dev/projects/Chen_environmental_study/R_code/data/real_data/NHANES/PCB_99_14/clean/individual/PCB_1999_2004_common.csv", header = T, stringsAsFactors = F) 
pre_cor <- cor(X_orignal)
p <- ncol(pre_cor)
set.seed(1234)
cor_h <- generate_chi(n = 10^5, p = p, pre_cor = pre_cor, structure = "un") %>% 
       add_inter(.) %>% cor(.)

X_test <- generate_chi(n = 150, p = p, pre_cor = pre_cor, structure = "un") %>% add_inter(.)   
cor(X_test) %>% offdiag(.) %>% hist(., main = "sample covariance n = 150")
cor(X_test) %>% offdiag(.) %>% summary(.)

par(mfrow=c(2,2))
X_decor1 <- true_value_method(input_data = X_test, emp = TRUE, combine = TRUE, sigma_total_emp = cor_h)$uncorr_data
cor_sample <- (X_decor1) %>% cor(.)
cor_sample %>% offdiag(.) %>% hist(., nclass = 40, main = "historical decor")
cor_sample %>% offdiag(.) %>% summary(.)

X_decor2 <- dgpGLASSO_method(X_decor1, rho = 0.1)$uncorr_data
cor_sample <- (X_decor2) %>% cor(.)
cor_sample %>% offdiag(.) %>% hist(., nclass = 40, main = "sparse decor 0.1")
cor_sample %>% offdiag(.) %>% summary(.)

X_decor2 <- dgpGLASSO_method(X_decor1, rho = 0.01)$uncorr_data
cor_sample <- (X_decor2) %>% cor(.)
cor_sample %>% offdiag(.) %>% hist(., nclass = 40, main = "sparse decor 0.01")
cor_sample %>% offdiag(.) %>% summary(.)

X_decor2 <- dgpGLASSO_method(X_decor1, rho = 0.005)$uncorr_data
cor_sample <- (X_decor2) %>% cor(.)
cor_sample %>% offdiag(.) %>% hist(., nclass = 40, main = "sparse decor 0.005")
cor_sample %>% offdiag(.) %>% summary(.)
```

### $X_h = XB, t = 8$
```{r}
## load the dateset 
library(R.utils)
library(MASS)
library(tidyverse)
library(foreach)
library(doRNG)
library(doParallel)
library(gtools) # for rbind based on columns
library(Matrix)
options(warn = 1, error = bettertrace::stacktrace)
setwd("~/dev/projects/Chen_environmental_study/")
sourceDirectory("./R_code/main_fn/",modifiedOnly = FALSE, recursive = TRUE)
sourceDirectory("./R_code/main_fn/method/",modifiedOnly = FALSE, recursive = TRUE)
source("./R_code/simulation_proposed_GCTA/local_helpers.R")
X_orignal <- read.csv("~/dev/projects/Chen_environmental_study/R_code/data/real_data/NHANES/PCB_99_14/clean/individual/PCB_1999_2004_common.csv", header = T, stringsAsFactors = F) 
pre_cor <- cor(X_orignal)
p <- ncol(pre_cor)
set.seed(1234)
cor_h <- generate_chi(n = 10^5, p = p, pre_cor = pre_cor, structure = "un") %>% tran_com(.) %>%
  add_inter(.) %>% cor(.)

X_test <- generate_chi(n = 150, p = p, pre_cor = pre_cor, structure = "un") %>% add_inter(.)   
cor(X_test) %>% offdiag(.) %>% hist(., main = "sample covariance n = 150")
cor(X_test) %>% offdiag(.) %>% summary(.)

par(mfrow=c(2,2))
X_decor1 <- true_value_method(input_data = X_test, emp = TRUE, combine = TRUE, sigma_total_emp = cor_h)$uncorr_data
cor_sample <- (X_decor1) %>% cor(.)
cor_sample %>% offdiag(.) %>% hist(., nclass = 40, main = "historical decor")
cor_sample %>% offdiag(.) %>% summary(.)

X_decor2 <- dgpGLASSO_method(X_decor1, rho = 0.1)$uncorr_data
cor_sample <- (X_decor2) %>% cor(.)
cor_sample %>% offdiag(.) %>% hist(., nclass = 40, main = "sparse decor 0.1")
cor_sample %>% offdiag(.) %>% summary(.)

X_decor2 <- dgpGLASSO_method(X_decor1, rho = 0.01)$uncorr_data
cor_sample <- (X_decor2) %>% cor(.)
cor_sample %>% offdiag(.) %>% hist(., nclass = 40, main = "sparse decor 0.01")
cor_sample %>% offdiag(.) %>% summary(.)

X_decor2 <- dgpGLASSO_method(X_decor1, rho = 0.005)$uncorr_data
cor_sample <- (X_decor2) %>% cor(.)
cor_sample %>% offdiag(.) %>% hist(., nclass = 40, main = "sparse decor 0.005")
cor_sample %>% offdiag(.) %>% summary(.)
```

## Simulation result on the variance estimation process 

### None
```{r}
result_path <- "~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/decor_method_None_sparse_method_None_result_list_fixed_sub_chi_structure_un_main_0.5_inter_0.1_n_100_200_500_1000_p_21_rho_e_0.5_decor_FALSE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_total_c_betam_8_c_betai_2_Var_cor"
file_list_all <- list.files(result_path,full.names = T)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect,
                             structure,
                             decor,
                             x_dist)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n]

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n]
summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

summary_final <- rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE) %>% setorder(., n)
additional
cbind(summary_final)

```
### Hist
```{r}
result_path <- "~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_sparse_decor_report_09_18_2019/tran_fn_tran_com_decor_method_hist_sparse_method__result_list_fixed_sub_chi_structure_un_main_0.5_inter_0.1_n_100_200_500_1000_p_21_rho_e_0.5_decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_total_c_betam_8_c_betai_2_Var_cor/"
file_list_all <- list.files(result_path,full.names = T)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect,
                             structure,
                             decor,
                             x_dist)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n]

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n]
summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

summary_final <- rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE) %>% setorder(., n)
additional
cbind(summary_final)

```

### Hist + Sparse(0.1)
```{r}
result_path <- "~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_sparse_decor_report_09_18_2019/tran_fn_tran_com_decor_hist_sparse_Glasso_grho_0.1_chi_structure_un_main_0.5_inter_0.1_n_100_200_500_1000_p_21_rho_e_0.5_decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_GCTA_est_total_c_betam_8_c_betai_2_Var_cor/"
file_list_all <- list.files(result_path,full.names = T)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect,
                             structure,
                             decor,
                             x_dist)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n]

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n]
summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

summary_final <- rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE) %>% setorder(., n)
additional
cbind(summary_final)

```


### Hist + Sparse(0.01)
```{r}
result_path <- "~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_sparse_decor_report_09_18_2019/tran_fn_tran_com_decor_hist_sparse_Glasso_grho_0.01_chi_structure_un_main_0.5_inter_0.1_n_100_200_500_1000_p_21_rho_e_0.5_decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_GCTA_est_total_c_betam_8_c_betai_2_Var_cor/"
file_list_all <- list.files(result_path,full.names = T)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect,
                             structure,
                             decor,
                             x_dist)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n]

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n]
summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

summary_final <- rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE) %>% setorder(., n)
additional
cbind(summary_final)
```


### Hist + Sparse(0.001)
```{r}
result_path <- "~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_sparse_decor_report_09_18_2019/tran_fn_tran_com_decor_hist_sparse_Glasso_grho_0.001_chi_structure_un_main_0.5_inter_0.1_n_100_200_500_1000_p_21_rho_e_0.5_decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_GCTA_est_total_c_betam_8_c_betai_2_Var_cor/"
file_list_all <- list.files(result_path,full.names = T)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect,
                             structure,
                             decor,
                             x_dist)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n]

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n]
summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

summary_final <- rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE) %>% setorder(., n)
additional
cbind(summary_final)
```

### Perfect Hist
```{r}
result_path <- "~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_sparse_decor_report_09_18_2019/tran_fn_null_tran_decor_hist_sparse__grho__chi_structure_un_main_0.5_inter_0.1_n_100_200_500_1000_p_21_rho_e_0.5_decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_GCTA_est_total_c_betam_8_c_betai_2_Var_cor/"
file_list_all <- list.files(result_path,full.names = T)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect,
                             structure,
                             decor,
                             x_dist)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n]

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n]
summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

summary_final <- rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE) %>% setorder(., n)
additional
cbind(summary_final)
```


### Perfect Hist + Sparse(0.1)
```{r}
result_path <- "~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_sparse_decor_report_09_18_2019/tran_fn_null_tran_decor_hist_sparse_Glasso_grho_0.1_chi_structure_un_main_0.5_inter_0.1_n_100_200_500_1000_p_21_rho_e_0.5_decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_GCTA_est_total_c_betam_8_c_betai_2_Var_cor/"
file_list_all <- list.files(result_path,full.names = T)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect,
                             structure,
                             decor,
                             x_dist)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n]

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n]
summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

summary_final <- rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE) %>% setorder(., n)
additional
cbind(summary_final)
```

### Perfect Hist + Sparse(0.01)
```{r}
result_path <- "~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_sparse_decor_report_09_18_2019/tran_fn_null_tran_decor_hist_sparse_Glasso_grho_0.01_chi_structure_un_main_0.5_inter_0.1_n_100_200_500_1000_p_21_rho_e_0.5_decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_GCTA_est_total_c_betam_8_c_betai_2_Var_cor/"
file_list_all <- list.files(result_path,full.names = T)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect,
                             structure,
                             decor,
                             x_dist)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n]

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n]
summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

summary_final <- rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE) %>% setorder(., n)
additional
cbind(summary_final)
```


### Perfect Hist + Sparse(0.001)
```{r}
result_path <- "~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_sparse_decor_report_09_18_2019/tran_fn_null_tran_decor_hist_sparse_Glasso_grho_0.001_chi_structure_un_main_0.5_inter_0.1_n_100_200_500_1000_p_21_rho_e_0.5_decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_GCTA_est_total_c_betam_8_c_betai_2_Var_cor/"
file_list_all <- list.files(result_path,full.names = T)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect,
                             structure,
                             decor,
                             x_dist)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n]

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n]
summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

summary_final <- rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE) %>% setorder(., n)
additional
cbind(summary_final)
```

<!-- # How to choose the tuning parameter for glasso -->

<!-- ```{r demo, include=F} -->
<!-- library(sas7bdat) -->
<!-- library(R.utils) -->
<!-- library(MASS) -->
<!-- library(tidyverse) -->
<!-- library(foreach) -->
<!-- library(doRNG) -->
<!-- library(doParallel) -->
<!-- library(gtools) # for rbind based on columns -->
<!-- options(warn = 1, error = bettertrace::stacktrace) -->
<!-- setwd("~/dev/projects/Chen_environmental_study/") -->
<!-- sourceDirectory("./R_code/main_fn/",modifiedOnly = FALSE, recursive = TRUE) -->
<!-- sourceDirectory("./R_code/main_fn/method/",modifiedOnly = FALSE, recursive = TRUE) -->
<!-- source("./R_code/simulation_proposed_GCTA/local_helpers.R") -->
<!-- # source("./reports/proposed_GCTA_paper/est_var_analysis/est_combined_data/covaraites_summary_2005_2014.R") -->
<!-- source("./reports/proposed_GCTA_paper/est_var_analysis/est_combined_data/covaraites_summary_1999_2004.R") -->

<!-- data <- PCB_1999_2004_common %>% std_fn(.) %>% add_inter(.) -->
<!-- hist_cov <- cov(data) -->
<!-- data_100 <- PCB_1999_2004_common[1:100, ] %>% true_value_method(input_data = ., ) -->

<!-- # data generation -->
<!-- emp_n <- nrow(PCB_1999_2004_common) -->
<!-- # n <- c(100,150,231) -->
<!-- n <- c(100,150,231) -->
<!-- dist <- "PCB" -->
<!-- generate_data <- generate_PCB -->
<!-- structure <- "un" -->
<!-- p <- length(PCB_common_1999)  -->
<!-- ``` -->