---
title: "Jackknife variance estimation corrections"
author: "Xuelong Wang"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    keep_tex: true
    fig_width: 12
    fig_height: 3
header-includes:
    - \usepackage{float,amsmath, bbm, siunitx, bm}
    - \usepackage{pdfpages}
    - \floatplacement{figure}{H}
    - \newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = "", message = FALSE, warning = FALSE, echo = FALSE, fig.height = 10)
options(scipen=1, digits=3)

library(MASS)
library(tidyverse)
library(data.table)
library(ggforce)
library(ggpubr)
library(gridExtra)
source("~/dev/projects/Chen_environmental_study/reports/proposed_GCTA_paper/est_var_analysis/est_combined_data/jackknife/simulation_result_jackknife_total_variance_correction.R")
```

# Motivation

Jackknife is one of sub-sampling technique to estimate the bias and variance of a statistics $S(X_1,\dots, X_n)$. Define $S_{(i)} = S(X_1,X_{(i-1)},X_{(i+1)} X_n)$, then the variance of $S$ estimated by jackknife is 
\[
Var(S) = \frac{n-1}{n}\sum_i(S_{(i)} - S_.)^2
\]
Where $S_. = \frac{1}{n}\sum_iS_{(i)}$.
Note that $\tilde{Var}(S(X_1, \dots, S_{n-1})) = \sum_i(S_{(i)} - S_.)^2$ can be conisder as a estimation for $Var(S(X_1, \dots, S_{n-1}))$
However, Eforn in 1981 shown that the jackknife variance esimtation is always overestimate the true variance. \[
  E(\tilde{Var}(S(X_1, \dots, S_{n-1}))) \geq Var(S(X_1, \dots, S_{n-1}))
\]

# One solution 

If we assume the $S$ is a smooth functions of emperical CDF, especially a quadratic functions, then it can be shown the leading terms of $E(\tilde{Var}(S(X_1, \dots, S_{n-1}))) \geq Var(S(X_1, \dots, S_{n-1}))$ is a quadratic term in expecation. Therefore we could try to estimate the quadratic term and correct the bias for the jackknife variance estimation.

Define $Q_{ii'} \equiv nS - (n-1)(S_{i} - S_{i'}) + (n-2)S_{(ii')}$, then the correction will be 
\[
\hat{Var}^{corr}(S(X_1, \dots, X_n)) = \hat{Var}(S(X_1, \dots, X_n)) - \frac{1}{n(n-1)}\sum_{i < i'}(Q_{ii'}- \bar{Q})^2
\]
where $\bar{Q} = \sum_{i < i'}(Q_{ii'})/(n(n-1)/2)$

# Simulation study 

## without correction

### setup 
- Independent 
- Normal
- p = 21
- with interaction terms

### simulation_result

```{r}
summary_final_none_21[,.(method, n, est_mean, est_var, var_jack)]
```

### setup 
- Independent 
- Normal
- p = 22
- with interaction terms

### simulation_result

```{r}
summary_final_none_22[,.(method, n, est_mean, est_var, var_jack)]
```
Note that based on the ideal situation the bias will reduced when sample size is large and distribution is normal.

## correction

### setup 
- Independent 
- Normal
- p = 21
- with interaction terms

### simulation result 

```{r}
additional_normal_total
summary_final_normal_total
```

### setup 
- Independent 
- Chi
- p = 21
- with interaction terms

### simulation result 

```{r}
additional_chi_total
summary_final_chi_total
```

### setup 
- PCB
- p = 21
- with interaction terms
- with decorrelation 

### simulation result 

```{r}
additional_PCB_total
summary_final_PCB_total
```


# Questions and other methods

## Questions
- Running time is large $n*(n-1)/2$
- Assumptions: quadratic form of S, $Var^n = \frac{n-1}{n}Var^{n-1}$? 
- The coefficient about the correction

## Other methods

# Some modification based on previous simulation

## Main effect only or larger n for combined effect

### Normal main n>p
```{r,}
additional_normal_main
summary_final_normal_main
additional_normal_main_rr
summary_final_normal_main_rr
```

### chi n >p
```{r,}
additional_chi_main
summary_final_chi_main
```

### Normal main n<p, p = 100
```{r,}
additional_normal_main_100
summary_final_normal_main_100
additional_normal_main_100_rr
summary_final_normal_main_100_rr
```

### Chi combined effect n <p p = 25
```{r,}
additional_chi_total_25
summary_final_chi_total_25
```

### PCB combined effec n<p p = 21
```{r,}
additional_PCB_total_corr
summary_final_PCB_total_corr
```

### PCB combined effec n<p p = 21 with rank transformation
```{r,}
additional_PCB_rank_each_total
summary_final_PCB_rank_each_total_corr
```

### PCB combined effec n<p p = 21 with rank transformation for all 
```{r,}
additional_PCB_rank_total
summary_final_PCB_rank_total_corr
```

