---
title: "Simulation of PCB with resampling"
author: "Xuelong Wang"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    keep_tex: true
    fig_width: 12
    fig_height: 6
header-includes:
    - \usepackage{float,amsmath, bbm, siunitx, bm}
    - \floatplacement{figure}{H}
    - \newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = FALSE, message = FALSE, warning = FALSE, fig.height = 10)
```



# Decorrelation using Graphical Lasso (glasso) and motivation 

Based on the previous simulation results, we found that the variance estimator of proposed method is not very stable, means variance of the estimator is large when the sample proportion is small. Thus, One possible reason for that is the decorrelation procedure of the proposed method is not that accurate. It uses the SVD method to find the inverse covariance matrix, which could be too closed to the sample date to capture the true structure of the covariance matrix. Therefore, we want to try the glasso method which add penalty to the estimated inverse covariance matrix if it's in favor of the sample data.

## Main Idea of glasso

For the N multivariate normal observations of dimension p, with mean $\mu$ and covaraince $\Sigma$. 

To estimate the precision matrix, we could maximize the penalized log-likelihood,

\[
  \log{|\Theta|} - tr(S\Theta) - \rho||\Theta||_1,
\]

where  

- $\Theta = \Sigma^{-1}$,   
- $||\cdot||_1$ is for $L_1$ norm - sum of all the absolute values of each elements of $\Theta$,  
- $S$ is the empirical covariance matrix.  

Note: as other lasso method, the key assumption of glasso is the sparsity of the covariance matrix $\Sigma$

# Simulation result

```{r load graph, include=FALSE}
source("./PCB_total_glasso_graph.R")
```

## Simulation result of the fixed fixed 

glasso

```{r fixed fixed total glasso}
for (i in 1:length(plot_PCB_fixed_fixed_total_glasso)){
  print(plot_PCB_fixed_fixed_total_glasso[[i]])
}

```

svd

```{r fixed fixed total svd}
plot_PCB_fixed_fixed_total_svd
```

# Conclusion

# Further work 
