---
title: "Simulation Fixed Effect"
author: "Xuelong Wang"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    keep_tex: true
    fig_width: 12
    fig_height: 6
header-includes:
    - \usepackage{float,amsmath, bbm, siunitx, bm}
    - \floatplacement{figure}{H}
    - \newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(data.table)
library(ggplot2)
knitr::opts_chunk$set(echo = FALSE)

data_clean_fn <- function(x) {
  y = fread(x); 
  y <- y[,-1]; 
  names(y)[1:6] <- c("true_main", "true_interaction", "GCTA_main", "GCTA_interaction", "prop_main", "prop_interaction")
  y_tmp <- tidyr::gather(data.frame(y[,1:6]));
  y_final <- cbind(y_tmp, y[,7:ncol(y)])
  }
```

# Motivation

Based on the previous simulation results, the interaction effect estimation is biased. One possible reason is that we consider the $\beta's$ as random effects, which will affect the covariance structure of the signals. More specifically, 
If $\beta$ is fixed, then we have following
\[
  Var(X\beta) = \beta^T\Sigma_{X}\beta,
\]
Where $\Sigma_{X} = Var(X)$. We can see that the X's covariance has an affect on the signal's variance. 

However, if we consider $\beta$ as an random effect, then $Var(X\beta)$ is not only related with $X$ but also with $\beta's$ covariance. Therefore, in the following simulation the $\beta$ is fixed and there is no average out summary across different values of $\beta$. 

# Simulation result 

## Setup
In the following simulation, I just randomly generated 3 different sets of values for $\beta$ (main and interaction) . Give each set of $\beta's$, the values of y is generated 100 times and thus 100 estimations of variance of main effects and interaction effects. 

## Full data

### Averaged estimation

```{r full data, results='asis'}
load("~/dev/projects/Chen_environmental_study/result/simulation_fixed_full_8tran")
for(i in 1:length(result_list)){
  
out_table <- result_list[[i]]
knitr::kable(x = out_table,
             format = "latex",
             caption = strsplit(names(result_list)[i], split = "_")[[1]][-1] %>% paste(.,collapse = "-"))  %>% 
  kableExtra::kable_styling(full_width = FALSE, 
                            latex_options = c("striped", "hold_position"), 
                            stripe_color = "gray!80")  %>%
  print(.)
}
```

\clearpage

### Histgram of 100 iterations

```{r Histgram full data}

input_file <- list.files("~/dev/projects/Chen_environmental_study/result/inter_result/")  %>% paste0("~/dev/projects/Chen_environmental_study/result/inter_result/",.)

file_path <- input_file[grepl(c("full_1"),input_file, perl=TRUE)]
input_table <- lapply(file_path, FUN = data_clean_fn ) %>% rbindlist(., fill = TRUE)
input_table[,("by") := as.character(input_table[["by"]])]
```

```{r full_main}
ggplot(data = input_table[grepl("main",key, perl=TRUE)], aes(x = key, y = value, fill = key)) +
  geom_violin(alpha = 0.2) +
  geom_boxplot(alpha = 0.7) +
  facet_wrap(facets = vars(tran,by,subset), ncol = 4, scales = "free") +
  ggtitle(paste0("Main effect full dataset")) 
  
```

```{r full_inter}
ggplot(data = input_table[grepl("interaction",key, perl=TRUE)], aes(x = key, y = value, fill = key)) +
  geom_violin(alpha = 0.2) +
  geom_boxplot(alpha = 0.7) +
  facet_wrap(facets = vars(tran,by,subset), ncol = 4, scales = "free") +
  ggtitle(paste0("Interaction effect full dataset")) 
```


## Remove 7 subset data

### Averaged estimation

```{r sub7 data, results='asis'}
load("~/dev/projects/Chen_environmental_study/result/simulation_fixed_sub7_8tran")
for(i in 1:length(result_list)){
  
out_table <- result_list[[i]]
knitr::kable(x = out_table,
             format = "latex",
             caption = strsplit(names(result_list)[i], split = "_")[[1]][-1] %>% paste(.,collapse = "-"))  %>% 
  kableExtra::kable_styling(full_width = FALSE, 
                            latex_options = c("striped", "hold_position"), 
                            stripe_color = "gray!80")  %>%
  print(.)
}
```

\clearpage

### Histgram of 100 iterations

```{r Histgram sub7 data}

input_file <- list.files("~/dev/projects/Chen_environmental_study/result/inter_result/")  %>% paste0("~/dev/projects/Chen_environmental_study/result/inter_result/",.)

file_path <- input_file[grepl(c("sub7_1"),input_file, perl=TRUE)]
input_table <- lapply(file_path, FUN = data_clean_fn ) %>% rbindlist(., fill = TRUE)
input_table[,("by") := as.character(input_table[["by"]])]
```

```{r sub7_main}
ggplot(data = input_table[grepl("main",key, perl=TRUE)], aes(x = key, y = value, fill = key)) +
  geom_violin(alpha = 0.2) +
  geom_boxplot(alpha = 0.7) +
  facet_wrap(facets = vars(tran,by,subset), ncol = 4, scales = "free") +
  ggtitle(paste0("Main effect sub7 dataset")) 
  
```

```{r sub7_inter}
ggplot(data = input_table[grepl("interaction",key, perl=TRUE)], aes(x = key, y = value, fill = key)) +
  geom_violin(alpha = 0.2) +
  geom_boxplot(alpha = 0.7) +
  facet_wrap(facets = vars(tran,by,subset), ncol = 4, scales = "free") +
  ggtitle(paste0("Interaction effect sub7 dataset")) 
```

## Remove 11 subset data

### Averaged estimation

```{r sub11 data, results='asis'}
load("~/dev/projects/Chen_environmental_study/result/simulation_fixed_sub11_8tran")
for(i in 1:length(result_list)){
  
out_table <- result_list[[i]]
knitr::kable(x = out_table,
             format = "latex",
             caption = strsplit(names(result_list)[i], split = "_")[[1]][-1] %>% paste(.,collapse = "-"))  %>% 
  kableExtra::kable_styling(full_width = FALSE, 
                            latex_options = c("striped", "hold_position"), 
                            stripe_color = "gray!80")  %>%
  print(.)
}
```

\clearpage

### Histgram of 100 iterations

```{r Histgram sub11 data}

input_file <- list.files("~/dev/projects/Chen_environmental_study/result/inter_result/")  %>% paste0("~/dev/projects/Chen_environmental_study/result/inter_result/",.)

file_path <- input_file[grepl(c("sub11_1"),input_file, perl=TRUE)]
input_table <- lapply(file_path, FUN = data_clean_fn ) %>% rbindlist(., fill = TRUE)
input_table[,("by") := as.character(input_table[["by"]])]
```

```{r sub11_main}
ggplot(data = input_table[grepl("main",key, perl=TRUE)], aes(x = key, y = value, fill = key)) +
  geom_violin(alpha = 0.2) +
  geom_boxplot(alpha = 0.7) +
  facet_wrap(facets = vars(tran,by,subset), ncol = 4, scales = "free") +
  ggtitle(paste0("Main effect sub11 dataset")) 
  
```

```{r sub11_inter}
ggplot(data = input_table[grepl("interaction",key, perl=TRUE)], aes(x = key, y = value, fill = key)) +
  geom_violin(alpha = 0.2) +
  geom_boxplot(alpha = 0.7) +
  facet_wrap(facets = vars(tran,by,subset), ncol = 4, scales = "free") +
  ggtitle(paste0("Interaction effect sub11 dataset")) 
```

# Conclusion

# Further work
