---
title: "Normal_cor_simulation"
author: "Xuelong Wang"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    keep_tex: true
    fig_width: 12
    fig_height: 10
header-includes:
    - \usepackage{float,amsmath, bbm, siunitx, bm}
    - \floatplacement{figure}{H}
    - \newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(data.table)
library(ggplot2)
knitr::opts_chunk$set(echo = FALSE)

# change to the long talble formate so that it can work with ggplot)(fill and gird)
data_clean_fn <- function(x) {
  y = fread(x); 
  y <- y[,-1]; 
  names(y)[1:6] <- c("true_main", "true_interaction", "GCTA_main", "GCTA_interaction", "prop_main", "prop_interaction")
  y_tmp <- tidyr::gather(data.frame(y[,1:6]));
  y_final <- cbind(y_tmp, y[,7:ncol(y)])
  }
```

# Motivation
Some theoretical results suggest that if the covariates follows normal distributions, then even they are not indepedent to each, we still can use the proposed method to estimate the interaction effect unbaisly.

## Model 
\[
  y_i = \alpha + \sum_{j = 1}^p x_j\beta_j + \sum_{j \neq k} \gamma_{ji}x_jx_k + \epsilon_i.
\]
Matrix form
\[
  y_i = X_i^T\beta + X_i^T\Gamma X_i + \epsilon_i,
\]
Where 

- $X_i = (x_{i1}, \dots x_{ip})^T$, in our case assumpe $X \sim N(0, \Sigma_p)$   
- $\epsilon \indep x_{ji}$  
- $\beta = (\beta_1, \dots, \beta_p)^T$ is fixed  
- $\Gamma$ is a $p \times p$ matrix with diagonal elements equal to 0.  



## Goal of the study 
\begin{align*}
Var(Y_i) &= Var(X_i^T\beta + X_i^T\Gamma X_i) + Var(\epsilon_i) \\
         &= Var(X_i^T\beta) + Var(X_i^T\Gamma X_i) + Cov(X_i^T\beta, X_i^T\Gamma X_i) + Var(\epsilon_i) \\
\end{align*}

\begin{align*}
Cov(X_i^T\beta, X_i^T\Gamma X_i) &= E[(X_i^T\beta - E(X_i^T\beta))(X_i^T\Gamma X_i - E(X_i^T\Gamma X_i))]\\
    &= E[X_i^T\beta(X_i^T\Gamma X_i - E(X_i^T\Gamma X_i)) \\
    &= E[X_i^T\beta(X_i^T\Gamma X_i - trace(\Gamma\Sigma_p))] \\
    &= E[X_i^T\beta \cdot X_i^T\Gamma X_i] \\
    &= E[(\sum_m(x_{im}\beta_m))(\sum_j\sum_k\gamma_{jk}x_{ij} x_{ik}) ] \\
    & = 0 \tag*{Note that $\gamma_{jj} = 0$}
\end{align*}


# Simulation result 

## Setup

\[
X \sim N(0, \Sigma_p)
\]
Where 
\[
\Sigma_p = \begin{bmatrix}
               1 & \dots  & \rho \\
               \vdots & \ddots & \vdots \\
               \rho & \dots  & 1
           \end{bmatrix}
\]
$\rho = \{0.1, \dots, 0.9 \}$

### Averaged estimation Correlated-Normal

```{r full data norm, results='asis'}
load("~/dev/projects/Chen_environmental_study/result/simulation_norm_core_null")
for(i in 1:length(result_list)){
  
out_table <- result_list[[i]]

out_table <- out_table[,c(1,3,5,2,4,6)]

knitr::kable(x = out_table,
             format = "latex",
             caption = gsub("_", "-", names(result_list)[i]))  %>% # replace _ by - b/c latex 
  kableExtra::add_header_above(., c("Main" = 3, "Interaction" = 3)) %>% 
  kableExtra::kable_styling(., full_width = FALSE,
                            latex_options = c("striped", "hold_position"),
                            stripe_color = "gray!80")  %>%
  print(.)
}
```

\clearpage

### Histgram of 100 iterations Correlated-Normal

```{r Histgram full data norm}

input_file <- list.files("~/dev/projects/Chen_environmental_study/result/inter_result/norm_cor/")  %>% paste0("~/dev/projects/Chen_environmental_study/result/inter_result/norm_cor/",.)

file_path <- input_file[grepl(c("null_1"),input_file, perl=TRUE)]
input_table <- lapply(file_path, FUN = data_clean_fn ) %>% rbindlist(., fill = TRUE)
input_table[,("cor") := as.character(input_table[["cor"]])] # chage the formate into character so that it can print out correctly
```

```{r normal_main}
ggplot(data = input_table[grepl("main",key, perl=TRUE)], aes(x = key, y = value, fill = key)) +
  geom_violin(alpha = 0.2) +
  geom_boxplot(alpha = 0.7) +
  facet_wrap(facets = vars(cor, tran), ncol = 3, scales = "free") +
  ggtitle(paste0("Main_effect_Normal_cor")) 
  
```

```{r normal_inter}
ggplot(data = input_table[grepl("interaction",key, perl=TRUE)], aes(x = key, y = value, fill = key)) +
  geom_violin(alpha = 0.2) +
  geom_boxplot(alpha = 0.7) +
  facet_wrap(facets = vars(cor, tran), ncol = 3, scales = "free") +
  ggtitle(paste0("Interactive_effect_Normal_cor")) 
```

\clearpage



### Averaged estimation Indepdent Chi-squared 

```{r full data chi, results='asis'}
load("~/dev/projects/Chen_environmental_study/result/simulation_chi_ind_null")
for(i in 1:length(result_list)){
  
out_table <- result_list[[i]]

out_table <- out_table[,c(1,3,5,2,4,6)]

knitr::kable(x = out_table,
             format = "latex",
             caption = gsub("_", "-", names(result_list)[i]))  %>% # replace _ by - b/c latex 
  kableExtra::add_header_above(., c("Main" = 3, "Interaction" = 3)) %>% 
  kableExtra::kable_styling(., full_width = FALSE,
                            latex_options = c("striped", "hold_position"),
                            stripe_color = "gray!80")  %>%
  print(.)
}
```

\clearpage

### Histgram of 100 iterations Indepdent Chi-squared

```{r Histgram full data chi}

input_file <- list.files("~/dev/projects/Chen_environmental_study/result/inter_result/chi_ind/")  %>% paste0("~/dev/projects/Chen_environmental_study/result/inter_result/chi_ind/",.)

file_path <- input_file[grepl(c("null_1"),input_file, perl=TRUE)]
input_table <- lapply(file_path, FUN = data_clean_fn ) %>% rbindlist(., fill = TRUE)
```

```{r chi_main}
ggplot(data = input_table[grepl("main",key, perl=TRUE)], aes(x = key, y = value, fill = key)) +
  geom_violin(alpha = 0.2) +
  geom_boxplot(alpha = 0.7) +
  ggtitle(paste0("Main_effect_Chi_ind")) 
  
```

```{r chi_inter}
ggplot(data = input_table[grepl("interaction",key, perl=TRUE)], aes(x = key, y = value, fill = key)) +
  geom_violin(alpha = 0.2) +
  geom_boxplot(alpha = 0.7) +
  ggtitle(paste0("Interactive_effect_Chi_ind")) 
```


# Conclusion

# Further work
