---
title: "Removing_non_normaly_covariates_simulation"
author: "Xuelong Wang"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_caption: yes
    fig_width: 20
    keep_md: yes
    number_sections: yes
    theme: united
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/dev/projects/Chen_environmental_study/")
options(width = 100)
library(ggplot2)
library(tidyverse)
library(sas7bdat)
library(printr)

#####################################################################################
## figures number function
#####################################################################################
# A function for captioning and referencing images
fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("Figure ", i, ": ", text, sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})

#####################################################################################
## figures number function
#####################################################################################
# A function for captioning and referencing tables

tab <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("Table ", i, ": ", text, sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})

#####################################################################################
## calculate MSE for the result 
#####################################################################################
sum_fn <- function(x) {
  x <- x[x[,1]!=0,]
  c(colMeans((x[,c(3,4)] -x[,c(1,2)])^2), colMeans((x[,c(5,6)] -x[,c(1,2)])^2))}

#####################################################################################
## plot function
#####################################################################################
plot_fn <- function(x, names) {
    x <- x[x[,1]!=0,]
    g1<- ggplot(tidyr::gather(data.frame(x[,c(1,3,5)])), aes(value, fill = key, y = ..density..)) +
    geom_histogram(alpha = 0.2, position = "identity", binwidth =  0.5) +
    geom_density(alpha = 0.2) +
    xlim(-20,20)+
    scale_fill_discrete(name="Methods",
                        labels=c("Original GCTA", "Proposed Method", "True value"))+
    theme(legend.title = element_text(size=25), plot.title = element_text(size = 30, face = "bold"))+
    ggtitle(paste0("main_total_effect_",names))
  
  
    g2 <- ggplot(tidyr::gather(data.frame(x[,c(2,4,6)])), aes(value, fill = key, y = ..density..)) +
    geom_histogram(alpha = 0.2, position = "identity", binwidth =  0.5) +
    geom_density(alpha = 0.2) +
    xlim(0,20)+
    scale_fill_discrete(name="Methods",
                        labels=c("Original GCTA", "Proposed Method", "True value"))+
    theme(legend.title = element_text(size=25), plot.title = element_text(size = 30, face = "bold"))+
    ggtitle(paste0("interaction_total_effect_",names))
    
    gridExtra::grid.arrange(g1, g2, ncol=2)
}

box_plot_fn <- function(x, names){
  x <- x[x[,1]!=0,]
  g1 <- ggplot(tidyr::gather(data.frame(x[,c(1,3,5)])), aes(x = key, y = value, color = key)) + 
    geom_boxplot() +
    ylim(-20,20)+
    labs(x = "method", y = "Variance")+
    scale_color_discrete(name="Methods",
                        labels=c("Original GCTA", "Proposed Method", "True value"))+
    theme(legend.title = element_text(size=25), plot.title = element_text(size = 30, face = "bold"))+
    ggtitle(paste0("main_total_effect_",names))
  
  g2 <- ggplot(tidyr::gather(data.frame(x[,c(2,4,6)])), aes(x = key, y = value, color = key)) + 
    geom_boxplot() +
    ylim(0, 20)+
    labs(x = "method", y = "Variance")+
    scale_color_discrete(name="Methods",
                        labels=c("Original GCTA", "Proposed Method", "True value"))+
    theme(legend.title = element_text(size=25), plot.title = element_text(size = 30, face = "bold"))+
    ggtitle(paste0("interaction_total_effect_",names))
    
  gridExtra::grid.arrange(g1, g2, ncol=2)
}
```

```{r loading result, include=FALSE}
source("./R_code/Yang_REML.R")
load("./result/simulation_rank_quantile_tranformation_non_normal_subset")
```

# Tranformation results 

The followings are the empirical density graphs of the covariates, which are selected based on their graphs shape. Those density graphs are not closed to a normal distribution

```{r tranformation result, echo = FALSE, fig.height=15, fig.width=15}

a=read.sas7bdat("~/dev/projects/Chen_environmental_study/R_code/pcbs1000nomiss.sas7bdat")
a=data.matrix(a[,2:35], rownames.force = NA)
b <- a 
b_rank <- std_fn(b, ncol(b), tran_FUN = rank_tran) 
b_quantile <- std_fn(b, ncol(b), tran_FUN = norm_quantile_tran) 

non_normal_index <- match(c("LBX157", "LBX167", "LBX189","LBX128", "LBX195", "LBX151", "LBX172", "LBX156", "LBX177", "LBX178", "LBX194"), colnames(a))
```

## Orignal scale

```{r original scale, echo=FALSE, fig.height=15, fig.width=15}
b <- b[,non_normal_index] 
melted_data <- reshape2::melt(b)
colnames(melted_data) <- c("ID", "PCB_var","value")
ggplot(melted_data, aes(x = value, y = ..density..)) +
  geom_histogram(alpha = 0.2, position = "identity", binwidth =  0.5) +
  geom_density(alpha = 0.2) +
  facet_wrap(PCB_var ~ ., nrow = 3, scales = "free_y")

```

## After the rank tranformation

```{r rank, echo=FALSE, fig.height=15, fig.width=15}
b_rank <- b_rank[,non_normal_index] 
melted_data <- reshape2::melt(b_rank)
colnames(melted_data) <- c("ID", "PCB_var","value")
ggplot(melted_data, aes(x = value, y = ..density..)) +
  geom_histogram(alpha = 0.2, position = "identity", binwidth =  0.5) +
  geom_density(alpha = 0.2) +
  facet_wrap(PCB_var ~ ., nrow = 3)
```

## After the normal quantile transformation
```{r qunatile, echo=FALSE, fig.height=15, fig.width=15}
b_quantile <- b_quantile[,non_normal_index] 
melted_data <- reshape2::melt(b_quantile)
colnames(melted_data) <- c("ID", "PCB_var","value")
ggplot(melted_data, aes(x = value, y = ..density..)) +
  geom_histogram(alpha = 0.2, position = "identity", binwidth =  0.5) +
  geom_density(alpha = 0.2) +
  facet_wrap(PCB_var ~ ., nrow = 3)
```


# Simulation result on the subset of PCB data

## rank tranformation

### On PCB (removing 7 variables)

```{r, echo=FALSE}
load("./result/simulation_rank_quantile_tranformation_non_normal_subset_7")
out_table <- t(sum_fn(result_list$PCB_rank))
rownames(out_table) <- "MSE"

knitr::kable(x = out_table, 
             row.names = TRUE,
             format = "html",
             caption = tab$cap("rank","MSE of estimated variance under rank transformation"))  %>% 
  kableExtra::kable_styling(full_width = FALSE)
```


```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("rank_box", "Box-plot of Main and Interaction variance with rank transformed data")}
  box_plot_fn(result_list$PCB_rank,"PCB_rank")
```

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("rank_hist", "Histogram of Main and Interaction variance with rank transformed data")}
  plot_fn(result_list$PCB_rank,"PCB_rank")
```

### On PCB (removing 11 variables)

```{r, echo=FALSE}
load("./result/simulation_rank_quantile_tranformation_non_normal_subset_11")
out_table <- t(sum_fn(result_list$PCB_rank))
rownames(out_table) <- "MSE"

knitr::kable(x = out_table, 
             row.names = TRUE,
             format = "html",
             caption = tab$cap("rank","MSE of estimated variance under rank transformation"))  %>% 
  kableExtra::kable_styling(full_width = FALSE)
```


```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("rank_box", "Box-plot of Main and Interaction variance with rank transformed data")}
  box_plot_fn(result_list$PCB_rank,"PCB_rank")
```

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("rank_hist", "Histogram of Main and Interaction variance with rank transformed data")}
  plot_fn(result_list$PCB_rank,"PCB_rank")
```

## Normal quantile tranformation with subset

### On PCB (removing 7 variables)

```{r, echo=FALSE}
load("./result/simulation_rank_quantile_tranformation_non_normal_subset_7")
out_table <- t(sum_fn(result_list$PCB_quantile))
rownames(out_table) <- "MSE"

knitr::kable(x = out_table, 
             row.names = TRUE,
             format = "html",
             caption = tab$cap("quantile","MSE of estimated variance under quantile transformation"))  %>% 
  kableExtra::kable_styling(full_width = FALSE)
```

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("quantile_transformation_box", "Box-plot of Main and Interaction variance with quantile transformed data")}
  box_plot_fn(result_list$PCB_quantile,"PCB_quantile")
```

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("quantile_transformation_hist", "Histogram of Main and Interaction variance with quantile transformed data")}
  plot_fn(result_list$PCB_quantile,"PCB_quantile")
```

### On PCB (removing 11 variables)

```{r, echo=FALSE}
load("./result/simulation_rank_quantile_tranformation_non_normal_subset_11")
out_table <- t(sum_fn(result_list$PCB_quantile))
rownames(out_table) <- "MSE"

knitr::kable(x = out_table, 
             row.names = TRUE,
             format = "html",
             caption = tab$cap("quantile","MSE of estimated variance under quantile transformation"))  %>% 
  kableExtra::kable_styling(full_width = FALSE)
```

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("quantile_transformation_box", "Box-plot of Main and Interaction variance with quantile transformed data")}
  box_plot_fn(result_list$PCB_quantile,"PCB_quantile")
```

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("quantile_transformation_hist", "Histogram of Main and Interaction variance with quantile transformed data")}
  plot_fn(result_list$PCB_quantile,"PCB_quantile")
```
