---
title: "Simulation result"
output: html_document
---
```{r, include=FALSE}
knitr::opts_knit$set(root.dir = "~/dev/projects/Chen_environmental_study/")
options(width = 100)
```

```{r, include=FALSE}
load("./result/simulation_result_reproduce")
```

# Reproduce the simulation result from the proposal
The following results basically mimic the results from the proposal 

## Simulation setting up

### PCB data with no intercation effect and no interaction in the model

#### simulation procedure

\[
  y_i = \sum^m_{j = 1}z_j*\beta^{(main)}_j ~+~ \epsilon_i
\]
where $\beta_j \sim N(0, 0.5)$ and $z_j$'s are the standardized covariates

#### model 

\[ 
  y_i = \mu + x_{i1}\beta_1, \dots, + x_im\beta_m + \epsilon_i
\]

Where m is the number of covariates.

#### Result

```{r,echo=FALSE}
result_list$PCB_no_inter_m
```

- second row is standard deviation


### PCB data with no intercation effect but with interaction in the model

#### simulation procedure

\[
  y_i = \sum^m_{j = 1}z_j*\beta^{(main)}_j + z_j~+~ \epsilon_i
\]
where $\beta_j \sim N(0, 0.5)$ and $z_j$'s are the standardized covariates

#### model 

\[ 
  y_i = \mu + x_{i1}\beta_1, \dots, + x_im\beta_m + x_{i1}x_{i2}\gamma_{12}, \dots, + x_{i(m-1)}x_{im}\gamma_{(m-1)m} + \epsilon_i
\]

#### Result

```{r,echo=FALSE}
result_list$PCB_no_inter
```

### PCB data with intercation effect AND with interaction in the model

#### simulation procedure

\[
  y_i = \sum^m_{j = 1}z_j*\beta^{(main)}_j + \sum^{m(m-1)/2}_{j=1}z^{(inter)}_j\beta^{(inter)}_j ~+~ \epsilon_i
\]
where $\beta^{(main)}_j \sim N(0, 0.5)$ and $z_j$'s are the standardized covariates
$\beta^{(inter)}_j \sim N(0, 0.1)$ and $z^{(inter)}_j$ are interaction terms of the standardized covariates


#### model 

\[ 
  y_i = \mu + x_{i1}\beta_1, \dots, + x_im\beta_m + x_{i1}x_{i2}\gamma_{12}, \dots, + x_{i(m-1)}x_{im}\gamma_{(m-1)m} + \epsilon_i
\]

#### Result

```{r,echo=FALSE}
result_list$PCB_with_inter
```

### Simulated Independent data with intercation effect AND with interaction in the model

#### simulation procedure

\[
  y_i = \sum^m_{j = 1}z_j*\beta^{(main)}_j + \sum^{m(m-1)/2}_{j=1}z^{(inter)}_j\beta^{(inter)}_j ~+~ \epsilon_i
\]

- $\beta^{(main)}_j \sim N(0, 0.5)$      
- $z_j$'s are the standardized covariates  
- The original covarites is $X \sim N(0, I_m)$  
- $\beta^{(inter)}_j \sim N(0, 0.1)$ and $z^{(inter)}_j$ are interaction terms of the standardized covariates  


#### model 

\[ 
  y_i = \mu + x_{i1}\beta_1, \dots, + x_im\beta_m + x_{i1}x_{i2}\gamma_{12}, \dots, + x_{i(m-1)}x_{im}\gamma_{(m-1)m} + \epsilon_i
\]

#### Result

```{r,echo=FALSE}
result_list$b_norm_uncorr
```

### Simulated correlated data with intercation effect AND with interaction in the model

#### simulation procedure

\[
  y_i = \sum^m_{j = 1}z_j*\beta^{(main)}_j + \sum^{m(m-1)/2}_{j=1}z^{(inter)}_j\beta^{(inter)}_j ~+~ \epsilon_i
\]

- $\beta^{(main)}_j \sim N(0, 0.5)$      
- $z_j$'s are the standardized covariates  
- The original covarites is $X \sim N(0, 2J_m+I_m)$  
- $\beta^{(inter)}_j \sim N(0, 0.1)$ and $z^{(inter)}_j$ are interaction terms of the standardized covariates  


#### model 

\[ 
  y_i = \mu + x_{i1}\beta_1, \dots, + x_im\beta_m + x_{i1}x_{i2}\gamma_{12}, \dots, + x_{i(m-1)}x_{im}\gamma_{(m-1)m} + \epsilon_i
\]

#### Result

```{r,echo=FALSE}
result_list$b_norm_corr
```


# Simulating different values for $\beta's$

## simulation setting up 
The simulation is almost same as previous ones, the only different part is that the true values of $\beta's$ are not fixed. In other words, both $\beta$ and $\epsilon$ are simulated more than once. 

More specifically,  
\[
  y_i = \sum^m_{j = 1}z_j*\beta^{(main)}_j + \epsilon_i
\]

- simulated 50 for each j, $\beta^{(main)}_j \sim N(0, 0.5)$
- given a set fixed values of $\beta^{(main)}_j$, simulated 5 values for $\epsilon_i$
- similar way is also applied to the interaction case

The reason to also generate simulating different values for $\beta$ is for getting an averaged evaluation of the performance of GCTA and proposed method. 



## result 

- Following is an example of the result
- Each TWO rows represents a different $\beta's$ value and its corresponding sample standard deviation  
- Note that the estimated standard deviation is large than previous simulation, one reason may be the smaller sample size (from 100 to 5). 

```{r, echo=FALSE}
rm(result_list)
load("./result/simulation_with_beta_rep")
head(result_list$PCB_no_inter_m)
```
### Summarized the result
We used MSE to summarized the performance of the across the different $\beta's$ value.

- following is the results for different setup
- If the simulation covariates are independent, uncorrelated covariates can reduce MSE a lot under some conditions. However, if they are dependent the proposed method cannot give a good estimation

#### MSE of estimated variance 
```{r, echo=FALSE}
sum_fn <- function(x) {
  x <- x[x[,1]!=0,]
  c(colMeans((x[,c(3,4)] -x[,c(1,2)])^2), colMeans((x[,c(5,6)] -x[,c(1,2)])^2))}
lapply(FUN = sum_fn, X = result_list)
```


# Another thought

There is another thought about how to estimate the variance when one considering the interaction terms. 

Basically, we combine the main effect and their two-way interaction as the covariates matrix T, and use T as the main effect's covariates to fit the model and the estimated main effect variance $\hat\sigma_{\beta} = \hat\sigma_{\beta} + \hat\sigma_{\gamma}$  

#### simulation procedure

The simulation procedure is same as before


#### model 

The only difference is treating both the main effect and interaction effects as main effect and uncorrlated together. 

\[ 
  y_i = \mu + t_{i1}\beta_1, \dots, + t_im\beta_m + t_{m+1}\beta_{m+1}, \dots, + t_{m+m(m-1)/2}\beta_{m+m(m-1)/2} + \epsilon_i
\]

#### MSE of the variance main effect
- Following is the MSE of the main effect by using the fitting method above
- From the theoriatical point of view, this may not be make sence.
- There is an improvement for the proposed method under the dependent normal covariates case. 

```{r, echo = FALSE}
rm(list = ls())
load("./result/GCTA_combined_main_and_interaction")
sum_fn <- function(x) {
  x <- x[x[,1]!=0,]
  x[,1] <- x[,1] + x[,2] # sum main and inter vairance
  c(colMeans((x[,c(3), drop = FALSE] - x[,1, drop = FALSE])^2), colMeans((x[,c(5), drop = FALSE] - x[,1, drop = FALSE])^2))}
lapply(FUN = sum_fn, X = result_list)

# plot_fn <- function(x) {
#   x <- x[x[,1]!=0,]
#   hist(x[,1], xlim=c(-10,10), col="red", breaks = 25)
#   hist(x[,2], add=TRUE, col=rgb(0, 1, 0, 0.5) )
# }
# 
# lapply(FUN = plot_fn, X = result_list)
```
