---
title: "PCB sub sampling simulation 2"
author: "Xuelong Wang"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    keep_tex: true
    fig_width: 12
    fig_height: 3
header-includes:
    - \usepackage{float,amsmath, bbm, siunitx, bm}
    - \usepackage{pdfpages}
    - \floatplacement{figure}{H}
    - \newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = "", message = FALSE, warning = FALSE, fig.height = 10)
options(scipen=1, digits=2)

library(MASS)
library(tidyverse)
require(gridExtra)

# heat map function for correlation matrxi

h_map <- function(x,cor = F) {
  if(cor == F){
    Cor <- cor(x) %>% abs(.)
  } else {
    Cor <- x %>% abs(.)
  }
  
  h1 <- ggplot(data = melt(Cor), aes(x=(Var1), y= (Var2), fill=value)) + 
    scale_fill_gradient(low = "white", high = "blue") +
    geom_tile() +
    theme(axis.text.x  = element_text(angle=90))
  h1
}
source("~/dev/projects/Chen_environmental_study/reports/proposed_GCTA_paper/est_var_analysis/est_combined_data/simulation_result_total_PCB_1999.R")
source("~/dev/projects/Chen_environmental_study/reports/proposed_GCTA_paper/est_var_analysis/est_combined_data/covaraites_summary_1999_2004.R")
```


# Motivation and issues
Based on the PCB data simulation results, if we don't do the standardization then there tends to have a larger bias for the estimated total effect than the standardized covariates. So for the following section, we try to figure out what are the exact reasons. 

The one feature and issues is that the values of each PCB is very small, which lead to small values in its covariance matrix. However, the correlation coefficients among those PCBs are very large. When I fit the model with those small values of PCBs we have found that the total variance estimation tends to have a bigger bias than I use the stanardized covariates. 

I suspected that the small values of the PCBs and their covariance matrix may affect the performance of the main and interaction variance estimation


# Standardization
- $Var(X^Ta) = 8$, rescale $a$ based on the $a^T\Sigma_{emp}a$
- $\Sigma_{emp}$ is calcuated by all the PCBs from 1999 with $n = 4*10^3$ 
In here the standardization means to normalization:
\[
  \tilde{Z} = D(X - \mu_x), 
\]
Where $D$ is the scaling matrix, $d_{11} = SD(X_1)$.
So after the standardization, the variance explained by $\tilde{Z}$ should be same as $X$, as following shows 
\[
Y = a_0 + X^Ta + \epsilon \Rightarrow Y = b_0 + \tilde{Z}^Tb + \epsilon
\]

\[
  Y = b_0 + \tilde{Z}^Tb + \epsilon = b_0 + (X^TDb - \mu_x^TDb) +\epsilon
\]
In order to have the same model, we only need take linear tranformation of the previous parameters
\[
   a = Db \text{ and } a_0 = b_0 - \mu_x^TDb.
\]
After the reparameterization, we found that the $Var(\tilde{Z}^Tb) = Var(X^Ta)$, so that it should not affect performance of variance estimation as following shows. 

# Simulation result for the main effect

## PCB main effect 


### original scale without decorrelation
```{r}
setwd("~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/")
result_path <- "decor_method_None_result_list_fixed_sub_PCB_structure_un_main_0.5_inter_0_n_100_150_231_500_1000_p_21_rho_e_0.5_dim_red_coeff__decor_FALSE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_main_year_1999_std_PCB_FALSE_c_betam_8_c_betai_2"
file_list_all <- list.files(paste0("./", result_path, "/")) %>% paste0(paste0("./", result_path, "/"),.)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)


additional <- sub_result[1,.(var_main_effect,
                             decor,
                             x_dist)]
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_main, na.rm = T),
                                      NA_i = mean(is.na(GCTA_main)),
                                      var_main_effect = var_main_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_main_effect))^2, na.rm = T),
                                                                                                est_var = var(est_mean, na.rm = T),
                                                                                                est_mean = mean(est_mean, na.rm = T),
                                                                                                NA_main = sum(NA_i)), by = n] %>%
                        setorder(., n)

summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]
cbind(additional, summary_final_GCTA)

```


### original scale decorrelation
```{r}
setwd("~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/")
result_path <- "decor_method_hist_result_list_fixed_sub_PCB_structure_un_main_0.5_inter_0_n_100_150_231_500_1000_p_21_rho_e_0.5_dim_red_coeff__decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_main_year_1999_std_PCB_FALSE_c_betam_8_c_betai_2"
file_list_all <- list.files(paste0("./", result_path, "/")) %>% paste0(paste0("./", result_path, "/"),.)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)


additional <- sub_result[1,.(var_main_effect,
                             decor,
                             x_dist)]
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_main, na.rm = T),
                                      NA_i = mean(is.na(GCTA_main)),
                                      var_main_effect = var_main_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_main_effect))^2, na.rm = T),
                                                                                                est_var = var(est_mean, na.rm = T),
                                                                                                est_mean = mean(est_mean, na.rm = T),
                                                                                                NA_main = sum(NA_i)), by = n] %>%
                        setorder(., n)

summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]
cbind(additional, summary_final_GCTA)
```

### std covariate without decorrelation
```{r}
setwd("~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/")
result_path <- "decor_method_None_result_list_fixed_sub_PCB_structure_un_main_0.5_inter_0_n_100_150_231_500_1000_p_21_rho_e_0.5_dim_red_coeff__decor_FALSE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_main_year_1999_std_PCB_TRUE_c_betam_8_c_betai_2"
file_list_all <- list.files(paste0("./", result_path, "/")) %>% paste0(paste0("./", result_path, "/"),.)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)


additional <- sub_result[1,.(var_main_effect,
                             decor,
                             x_dist)]
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_main, na.rm = T),
                                      NA_i = mean(is.na(GCTA_main)),
                                      var_main_effect = var_main_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_main_effect))^2, na.rm = T),
                                                                                                est_var = var(est_mean, na.rm = T),
                                                                                                est_mean = mean(est_mean, na.rm = T),
                                                                                                NA_main = sum(NA_i)), by = n] %>%
                        setorder(., n)

summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]
cbind(additional, summary_final_GCTA)
```

### std covariate decorrelation
```{r}
setwd("~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/")
result_path <- "decor_method_hist_result_list_fixed_sub_PCB_structure_un_main_0.5_inter_0_n_100_150_231_500_1000_p_21_rho_e_0.5_dim_red_coeff__decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_main_year_1999_std_PCB_TRUE_c_betam_8_c_betai_2"
file_list_all <- list.files(paste0("./", result_path, "/")) %>% paste0(paste0("./", result_path, "/"),.)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)


additional <- sub_result[1,.(var_main_effect,
                             decor,
                             x_dist)]
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_main, na.rm = T),
                                      NA_i = mean(is.na(GCTA_main)),
                                      var_main_effect = var_main_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_main_effect))^2, na.rm = T),
                                                                                                est_var = var(est_mean, na.rm = T),
                                                                                                est_mean = mean(est_mean, na.rm = T),
                                                                                                NA_main = sum(NA_i)), by = n] %>%
                        setorder(., n)

summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]
cbind(additional, summary_final_GCTA)
```

\newpage

## PCB total effect

After adding the interaction terms, the difference between original and standardized scores are different. 

### original scale without decorrelation
```{r}
setwd("~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/")
result_path <- "decor_method_None_result_list_fixed_sub_PCB_structure_un_main_0.5_inter_0.1_n_100_150_231_500_1000_p_21_rho_e_0.5_dim_red_coeff__decor_FALSE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_total_year_1999_std_PCB_FALSE_c_betam_8_c_betai_2"
file_list_all <- list.files(paste0("./", result_path, "/")) %>% paste0(paste0("./", result_path, "/"),.)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)


# additional <- sub_result[1,.(total = var_total_effect,
#                              decor,
#                              x_dist)]

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n] %>% setorder(.,n)

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n] %>% setorder(.,n)

summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]
additional
rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE)
```


### original scale decorrelation
```{r}
setwd("~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/")
result_path <- "decor_method_hist_result_list_fixed_sub_PCB_structure_un_main_0.5_inter_0.1_n_100_150_231_500_1000_p_21_rho_e_0.5_dim_red_coeff__decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_total_year_1999_std_PCB_FALSE_c_betam_8_c_betai_2"
file_list_all <- list.files(paste0("./", result_path, "/")) %>% paste0(paste0("./", result_path, "/"),.)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)


# additional <- sub_result[1,.(total = var_total_effect,
#                              decor,
#                              x_dist)]

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n] %>% setorder(.,n)

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n] %>% setorder(.,n)

summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]
additional
rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE)
```

### std without decorrelation
```{r}
setwd("~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/")
result_path <- "decor_method_None_result_list_fixed_sub_PCB_structure_un_main_0.5_inter_0.1_n_100_150_231_500_1000_p_21_rho_e_0.5_dim_red_coeff__decor_FALSE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_total_year_1999_std_PCB_TRUE_c_betam_8_c_betai_2"
file_list_all <- list.files(paste0("./", result_path, "/")) %>% paste0(paste0("./", result_path, "/"),.)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)


# additional <- sub_result[1,.(total = var_total_effect,
#                              decor,
#                              x_dist)]

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n] %>% setorder(.,n)

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n] %>% setorder(.,n)

summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

additional
rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE)
```

### original scale decorrelation
```{r}
setwd("~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/")
result_path <- "decor_method_hist_result_list_fixed_sub_PCB_structure_un_main_0.5_inter_0.1_n_100_150_231_500_1000_p_21_rho_e_0.5_dim_red_coeff__decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_total_year_1999_std_PCB_TRUE_c_betam_8_c_betai_2"
file_list_all <- list.files(paste0("./", result_path, "/")) %>% paste0(paste0("./", result_path, "/"),.)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)


# additional <- sub_result[1,.(total = var_total_effect,
#                              decor,
#                              x_dist)]

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n] %>% setorder(.,n)

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n] %>% setorder(.,n)

summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]
additional
rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE)
```

# Simulation study on the small small values



## Normal 

### original scale without decorrelation

```{r}
setwd("~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/")
result_path <- "decor_method_None_result_list_fixed_sub_normal_structure_un_main_0.5_inter_0.1_n_100_150_231_500_1000_p_21_rho_e_0.5_dim_red_coeff__decor_FALSE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_total_c_betam_8_c_betai_2_cov"
file_list_all <- list.files(paste0("./", result_path, "/")) %>% paste0(paste0("./", result_path, "/"),.)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)


# additional <- sub_result[1,.(total = var_total_effect,
#                              decor,
#                              x_dist)]

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n] %>% setorder(.,n)

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n] %>% setorder(.,n)

summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

additional
rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE)
```



### original scale with decorrelation
```{r}
setwd("~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/")
result_path <- "decor_method_hist_result_list_fixed_sub_normal_structure_un_main_0.5_inter_0.1_n_100_150_231_500_1000_p_21_rho_e_0.5_dim_red_coeff__decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_total_c_betam_8_c_betai_2_cov"
file_list_all <- list.files(paste0("./", result_path, "/")) %>% paste0(paste0("./", result_path, "/"),.)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)


# additional <- sub_result[1,.(total = var_total_effect,
#                              decor,
#                              x_dist)]

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n] %>% setorder(.,n)

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n] %>% setorder(.,n)

summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]
additional
rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE)
```




## Chi 

### original scale without decorrelation

```{r}
setwd("~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/")
result_path <- "decor_method_None_result_list_fixed_sub_chi_structure_un_main_0.5_inter_0.1_n_100_150_231_500_1000_p_21_rho_e_0.5_dim_red_coeff__decor_FALSE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_total_c_betam_8_c_betai_2_cov"
file_list_all <- list.files(paste0("./", result_path, "/")) %>% paste0(paste0("./", result_path, "/"),.)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)


# additional <- sub_result[1,.(total = var_total_effect,
#                              decor,
#                              x_dist)]

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n] %>% setorder(.,n)

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n] %>% setorder(.,n)

summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]

additional
rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE)
```



### original scale with decorrelation
```{r}
setwd("~/dev/projects/Chen_environmental_study/result/simulation_proposed_GCTA_paper/var_est/combined_effects_PCBs_report_08_30_2019/")
result_path <- "decor_method_hist_result_list_fixed_sub_chi_structure_un_main_0.5_inter_0.1_n_100_150_231_500_1000_p_21_rho_e_0.5_dim_red_coeff__decor_TRUE_subpro_0_iter_100_nsub_1_EigenPrism_kernel_GCTA_kernel_est_total_c_betam_8_c_betai_2_cov"
file_list_all <- list.files(paste0("./", result_path, "/")) %>% paste0(paste0("./", result_path, "/"),.)
file_list <- file_list_all
sub_result <- lapply(file_list, function (x) {read.csv(x, header = TRUE, stringsAsFactors = FALSE)}) %>% rbindlist(., fill = TRUE)


# additional <- sub_result[1,.(total = var_total_effect,
#                              decor,
#                              x_dist)]

additional <- sub_result[1,.(var_main_effect,
                             var_inter_effect,
                             cov_main_inter_effect,
                             var_total_effect)]


# replace the 999 as NA 
sub_result[sub_EigenPrism_total == 999, c("sub_EigenPrism_total", "sub_EigenPrism_CI1", "sub_EigenPrism_CI2") := list(NA,NA,NA)]
sub_result[EigenPrism_total == 999, c("EigenPrism_total", "EigenPrism_CI1", "EigenPrism_CI2") := list(NA,NA,NA)]


# EigenPrism
summary_result_Eg <- sub_result[, .(est_mean = mean(EigenPrism_total, na.rm = T),
                                    NA_i = mean(is.na(EigenPrism_total)),
                                    var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                            est_var = var(est_mean, na.rm = T),
                                                                                            est_mean = mean(est_mean, na.rm = T),
                                                                                            NA_total = sum(NA_i)), by = n] %>% setorder(.,n)

summary_final_Eg <- cbind(summary_result_Eg)
summary_final_Eg[,method := "EigenPrism"]

# GCTA
summary_result_GCTA <- sub_result[, .(est_mean = mean(GCTA_total, na.rm = T),
                                      NA_i = mean(is.na(GCTA_total)),
                                      var_total_effect = var_total_effect[1]), by = .(i,n)][, .(MSE = mean((est_mean - mean(var_total_effect))^2, na.rm = T),
                                                                                              est_var = var(est_mean, na.rm = T),
                                                                                              est_mean = mean(est_mean, na.rm = T),
                                                                                              NA_total = sum(NA_i)), by = n] %>% setorder(.,n)

summary_final_GCTA <- cbind(summary_result_GCTA)
summary_final_GCTA[,method:= "GCTA"]
additional
rbindlist(list(summary_final_Eg, summary_final_GCTA), fill = TRUE)
```

