---
title: "PCB data validation (before 2004 and after 2004)"
author: "Xuelong Wang"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    keep_tex: true
    fig_width: 12
    fig_height: 6
header-includes:
    - \usepackage{float,amsmath, bbm, siunitx, bm}
    - \floatplacement{figure}{H}
    - \newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = FALSE, message = FALSE, warning = FALSE, fig.height = 10)
```

# Motivation
To get a better model performance evaluation of the proposed method, we need make sure that the sample size $n$ is greater than the combined covaraites(main and interaction) $p$. 
Therefore, we want to add more observations to our simuation dataset. But the format of the PCB data has changed a little after 2004 in the nhanes database. So we need to run some basic data validation steps to make sure that we can combine those data from different years together.

```{r loading the graph}
source("~/dev/projects/Chen_environmental_study/R_code/data/data_validation.R")
```

# Comparison of the 2000 and 2005 PCB data

## sample size

PCB_2000 `r nrow(pcb_chen)`
PCB_2005 `r nrow(pcb_05_06_no_missing)`

## box-plot comparison for each covairates
```{page 1}
compare_Chen_pcb_05_06[[1]]
```

```{page 2}
compare_Chen_pcb_05_06[[2]]
```

```{page 2}
compare_Chen_pcb_05_06[[2]]
```

```{page 2}
compare_Chen_pcb_05_06[[2]]
```