---
title: "Transformation_simulation"
author: "Xuelong Wang"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_caption: yes
    fig_width: 18
    keep_md: yes
    number_sections: yes
    theme: united
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/dev/projects/Chen_environmental_study/")
options(width = 100)
library(ggplot2)
library(tidyverse)
library(sas7bdat)
library(printr)

#####################################################################################
## figures number function
#####################################################################################
# A function for captioning and referencing images
fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("Figure ", i, ": ", text, sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})

#####################################################################################
## figures number function
#####################################################################################
# A function for captioning and referencing tables

tab <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("Table ", i, ": ", text, sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})

#####################################################################################
## calculate MSE for the result 
#####################################################################################
sum_fn <- function(x) {
  x <- x[x[,1]!=0,]
  c(colMeans((x[,c(3,4)] -x[,c(1,2)])^2), colMeans((x[,c(5,6)] -x[,c(1,2)])^2))}

#####################################################################################
## plot function
#####################################################################################
plot_fn <- function(x, names) {
    x <- x[x[,1]!=0,]
    g1<- ggplot(tidyr::gather(data.frame(x[,c(1,3,5)])), aes(value, fill = key, y = ..density..)) +
    geom_histogram(alpha = 0.2, position = "identity", binwidth =  0.5) +
    geom_density(alpha = 0.2) +
    xlim(-30,50)+
    labs(fill="Methods")+
    ggtitle(paste0("main_total_effect_",names))
  
    g2 <- ggplot(tidyr::gather(data.frame(x[,c(2,4,6)])), aes(value, fill = key, y = ..density..)) +
    geom_histogram(alpha = 0.2, position = "identity", binwidth =  0.5) +
    geom_density(alpha = 0.2) +
    xlim(0,40)+
    labs(fill="Methods")+
    ggtitle(paste0("interaction_total_effect_",names))
    
    gridExtra::grid.arrange(g1, g2, ncol=2)
}

box_plot_fn <- function(x, names){
  x <- x[x[,1]!=0,]
  g1 <- ggplot(tidyr::gather(data.frame(x[,c(1,3,5)])), aes(x = key, y = value, color = key)) + 
    geom_boxplot() +
    ylim(-30,50)+
    labs(color ="Methods")+
    labs(x = "method", y = "Variance")+
    ggtitle(paste0("main_total_effect_",names))
  
  g2 <- ggplot(tidyr::gather(data.frame(x[,c(2,4,6)])), aes(x = key, y = value, color = key)) + 
    geom_boxplot() +
    ylim(0, 40)+
    labs(color ="Methods")+
    labs(x = "method", y = "Variance")+
    ggtitle(paste0("interaction_total_effect_",names))
    
  gridExtra::grid.arrange(g1, g2, ncol=2)
}
```

```{r loading result, include=FALSE}
source("./R_code/Yang_REML.R")
load("./result/simulation_log_sqrt_tranformation")
```

# Transformation methods

- Based the density and histogram plots, the most of the empirical distributions of individual PCB covariate are __right__ skewed (heavy tail)
- I have tried 5 different transformation methods  `square root`, `log`, `box-cox`, `rank` and `quantile`
- Transformation is made __BEFORE__ generating the main and interaction signals

# Simulation and Models
\[
  y_i = \sum^m_{j = 1}z_j\beta^{(main)}_j + \sum^{m(m-1)/2}_{k=2}z^{(inter)}_k\beta^{(inter)}_k ~+~ \epsilon_i
\]

- Where $\beta^{(main)} \sim N(0, 0.5)$ and $\beta^{(intercation)} \sim N(0,0.1)$ are random effects
- During the simulation, the interaction coefficients are not zero and generated from a normal distribution
- For main and interaction effects ($\beta^{(main)}$ and $\beta^{(inter)}$), 80 groups of values are randomly generated
- Given a fixed main and interaction effect, responses are generated 20 times in order to get a averaged estimators of main and interaction variance
- For all of the following results, the model is same. 

# Steps
1. All 5 transformation methods are applied to the PCB data directly
2. Standardizing the transformed data into Z, where $E(Z_j) = 0, Var(Z_i) = 1$
3. For the proposed method, the input data Z is also transformed into __Uncorrelated__ by SVD method
4. Using the RMEL method to estimated the variances of main effect and interact effect


# Simulation Result for each method

- The following tables are about the MSE of estimated variance (averaging across different values of $\beta's$ )
- Histogram plots are about the empirical distribution of the true variances, GCTA estimator and Proposed estimator


## No transformation
- In this case, no transformation is applied to PCB data. This could be considered as a reference result

```{r, echo=FALSE}
out_table <- t(sum_fn(result_list$PCB))
rownames(out_table) <- "MSE"

knitr::kable(x = out_table, 
             row.names = TRUE,
             format = "html",
             caption = tab$cap("original scale","MSE of estimated varaince under original scale"))  %>% 
  kableExtra::kable_styling(full_width = FALSE)
```
- In the Table `r tab$ref("original scale")`, "GCTA_main" and "GCTA_interaction" stands for the estimation of GCTA method. 
- "pro_main" and "pro_interaction" stand for the estimation of proposed method. 
- MSE is calculated by $\sum(\sigma^2_{estimation} - \sigma^2_{true})^2/n$
- n in our case is 80

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("Original scale_box", "Box-plot of Main and Interaction variance with original scale data")}
  box_plot_fn(result_list$PCB,"PCB")
```

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("Original scale", "Histogram of Main and Interaction variance with original scale data")}
  plot_fn(result_list$PCB,"PCB")
```

- In figure `r fig$ref("Original scale")`, left hand-side is the histogram of main effect variance and right hand-side is for interaction effect variance. 
- GCTA_main (in pink) represents the estimated main effect's variance by GCTA method 
- pro_main (in green) represents the estimated main effect's variance by proposed method which is GCTA method with uncorrelated procedure
- True_main (in blue) represents the real variance of the main effect  
- It seems that the there is a bias for both GCTA and proposed estimators of interaction effect

## Square root transformation

- In this case, square root transformation is applied to PCB data. 


```{r, echo=FALSE}
out_table <- t(sum_fn(result_list$PCB_sqrt))
rownames(out_table) <- "MSE"

knitr::kable(x = out_table, 
             row.names = TRUE,
             format = "html",
             caption = tab$cap("sqrt","MSE of estimated variance under square root transformation"))  %>% 
  kableExtra::kable_styling(full_width = FALSE)
```

- In the Table `r tab$ref("sqrt")`, MSE is reduced by using square root transformation but not as much as log or cox-box transformation

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("sqrt_box", "Box-plot of Main and Interaction variance with square root transformed data")}
  box_plot_fn(result_list$PCB_sqrt,"PCB_sqrt")
```

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("sqrt", "Histogram of Main and Interaction variance with square root transformed data")}
  plot_fn(result_list$PCB_sqrt,"PCB_sqrt")
```

- In the figure `r fig$ref("sqrt")`, the histogram is similar with the original scale data's result


## Log transformation
- In this case, $log$ transformation applied to each column of PCB data.

```{r, echo=FALSE}
out_table <- t(sum_fn(result_list$PCB_log))
rownames(out_table) <- "MSE"

knitr::kable(x = out_table, 
             row.names = TRUE,
             format = "html",
             caption = tab$cap("log","MSE of estimated variance under log transformation"))  %>% 
  kableExtra::kable_styling(full_width = FALSE)
```

- In the Table `r tab$ref("log")`, Log transformation reduce the MSE a lot
- However, at least based on this simulation result GCTA method's result is better than proposed method

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("log", "Box-plot of Main and Interaction variance with log transformed data")}
  box_plot_fn(result_list$PCB_log,"PCB_log")
```

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("log", "Histogram of Main and Interaction variance with log transformed data")}
  plot_fn(result_list$PCB_log,"PCB_log")
```

- In figure `r fig$ref("log")`,after transformation both methods can predict the main effect well

- But for the interaction effect, the proposed method seems to have a bias which I will make a further study.


## Cox-Box transformation
- In this case, Cox-Box transformation is applied to each column of PCB data.
\[
 y(\lambda)=\begin{cases}
               \frac{y^\lambda - 1}{\lambda}, \lambda \neq 0\\
               log{(y)} , \lambda = 0
            \end{cases}
\]
- The choosing of $\lambda$ is based on a linear model $y = \beta_0$, where $\beta_0$ is the intercept.


```{r, echo=FALSE}
out_table <- t(sum_fn(result_list$PCB_cox))
rownames(out_table) <- "MSE"

knitr::kable(x = out_table, 
             row.names = TRUE,
             format = "html",
             caption = tab$cap("cox","MSE of estimated variance under cox-box transformation"))  %>% 
  kableExtra::kable_styling(full_width = FALSE)
```

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("cox_box", "Box-plot of Main and Interaction variance with cox-box transformed data")}
  box_plot_fn(result_list$PCB_cox,"PCB_cox")
```

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("cox", "Histogram of Main and Interaction variance with cox-box transformed data")}
  plot_fn(result_list$PCB_cox,"PCB_cox")
```

- The result is similar with log transformation's result




## Rank tranformation

```{r, include = FALSE}
load("./result/simulation_rank_quantile_tranformation")
```
- In this case, the ranks of each covariate are used to replace real values.

```{r, echo=FALSE}
out_table <- t(sum_fn(result_list$PCB_rank))
rownames(out_table) <- "MSE"

knitr::kable(x = out_table, 
             row.names = TRUE,
             format = "html",
             caption = tab$cap("rank","MSE of estimated variance under rank transformation"))  %>% 
  kableExtra::kable_styling(full_width = FALSE)
```
- In the Table `r tab$ref("rank")`, MSE is reduced by using rank transformation. 
- Compared with other methods, Using rank transformation seems to give the __BEST__ result in term of MSE


```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("rank", "Box-plot of Main and Interaction variance with rank transformed data")}
  box_plot_fn(result_list$PCB_rank,"PCB_rank")
```

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("rank", "Histogram of Main and Interaction variance with rank transformed data")}
  plot_fn(result_list$PCB_rank,"PCB_rank")
```

- In the figure `r fig$ref("rank")`, for the main effect, both GCTA and proposed methods are able to give a good estimations for the signals' variance. 
- However, for the interaction effect, the proposed method seems to have a bias in estimating the interaction variance. 

## Normal Quantile tranformation
- In this case, normal quantile are used the replace the original value for each variable of PCB data

* the Normal quantile Procedure is 
    1. Calculate the empirical precentails for each column of PCB data 
    1. Using the precentails to find the corresponding quantile value based on a Standard Normal distribution.
    1. Using the normal quantiles to replace the real values of PCB data column by column 


```{r, echo=FALSE}
out_table <- t(sum_fn(result_list$PCB_quantile))
rownames(out_table) <- "MSE"

knitr::kable(x = out_table, 
             row.names = TRUE,
             format = "html",
             caption = tab$cap("quantile","MSE of estimated variance under quantile transformation"))  %>% 
  kableExtra::kable_styling(full_width = FALSE)
```

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("quantile", "Box-plot of Main and Interaction variance with quantile transformed data")}
  box_plot_fn(result_list$PCB_quantile,"PCB_quantile")
```

```{r , echo = FALSE, warning=FALSE, fig.cap=fig$cap("quantile", "Histogram of Main and Interaction variance with quantile transformed data")}
  plot_fn(result_list$PCB_quantile,"PCB_quantile")
```