---
title: "Jackknife variance estimation corrections"
author: "Xuelong Wang"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    keep_tex: true
    fig_width: 12
    fig_height: 3
header-includes:
    - \usepackage{float,amsmath, bbm, siunitx, bm}
    - \usepackage{pdfpages}
    - \floatplacement{figure}{H}
    - \newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = "", message = FALSE, warning = FALSE, echo = FALSE, fig.height = 10)
options(scipen=1, digits=3)

library(MASS)
library(tidyverse)
library(data.table)
library(ggforce)
library(ggpubr)
library(gridExtra)
```

# Jackknife variance correction 

If we assume the $S$ is a smooth functions of emperical CDF, especially a quadratic functions, then it can be shown the leading terms of $E(\tilde{Var}(S(X_1, \dots, S_{n-1}))) \geq Var(S(X_1, \dots, S_{n-1}))$ is a quadratic term in expecation. Therefore we could try to estimate the quadratic term and correct the bias for the jackknife variance estimation.

Define $Q_{ii'} \equiv nS - (n-1)(S_{i} + S_{i'}) + (n-2)S_{(ii')}$, then the correction will be 
\[
\hat{Var}^{corr}(S(X_1, \dots, X_n)) = \hat{Var}(S(X_1, \dots, X_n)) - \frac{1}{n(n-1)}\sum_{i < i'}(Q_{ii'}- \bar{Q})^2
\]
where $\bar{Q} = \sum_{i < i'}(Q_{ii'})/(n(n-1)/2)$

# Simulation study compare two GCTA and GCTA_rr
GCTA_rr is the `mixed.solve` function from `rrBLUP` r package.  
Based on the following simulation results, 

1. when $n<p$ case, those two methods' results are very closed to each other. 
1. when $n>p$ case, in terms of effect estimation and jackknife variance estimation those two methods's reuslts are similar to each other. But for the variance corrections are quite different. That is the statistics $Q$ of our method has a very large variance which leads to negative correction result.  

### setup 
- Independent 
- Normal
- $p = 100$
- $n = \{50, 75,100, 150, 200\}$
- with interaction terms
- main effect: $Var(X^T\beta) = 8$ 
\newpage

```{r, include=F}
source("~/dev/projects/Chen_environmental_study/reports/proposed_GCTA_paper/est_var_analysis/est_combined_data/jackknife/GCTA_and_rr_v_jack_correction.R")
```

### n = 50
```{r}
res_50
h_50
```
\newpage

### n = 75
```{r}
res_75
h_75
```
\newpage

### n = 100
```{r}
res_100
h_100
```
\newpage

### n = 150
```{r}
res_150
h_150
```
\newpage

### n = 200
```{r}
res_200
h_200
```

### setup 
- Independent 
- Normal
- $p = 100$
- $n = \{50, 75,100, 150, 200\}$
- with interaction terms
- main effect: $Var(X^T\beta) = 0$ 
\newpage

```{r, include=F}
source("~/dev/projects/Chen_environmental_study/reports/proposed_GCTA_paper/est_var_analysis/est_combined_data/jackknife/GCTA_and_rr_v_jack_correction_main_0.R")
```

### n = 50
```{r}
res_50
h_50
```
\newpage

### n = 75
```{r}
res_75
h_75
```
\newpage

### n = 100
```{r}
res_100
h_100
```
\newpage

### n = 150
```{r}
res_150
h_150
```
\newpage

### n = 200
```{r}
res_200
h_200
```

## With iterations

### setup
