## Jackknife variance estimation on high dimension signal estimation
Different methods have their own 

### Jackknife variance estimation's bias and sample size n 

#### setup 
- Independent 
- Normal
- $p = \{100,1000\}$
- $n = \{50, 75,100, 150, 200, 500, 750,1000, 1500\}$
- $d = \{0.5, 0.75\} \times n$ or $d = 25$
- $n_{repeat} = n^{1.5}$ for delete d jackknife and $n_{repeat} = n$ for delete 1 jackknife 
- main effect: $Var(X^T\beta) = 8$ 

#### result
Based on the previous simulation results, we find there is a bias among all the jackknife variance estimation. Based on the Efron's result, the overestimation is because the statistics $S$ is not a smooth function of the distribution function, so that the correct coefficient actually inflate the variance estimation. 

The following reuslt is trying to see the relation between the bias and the sample size n 

Note:
1. For delete-1 jackknife, the variance estimation becomes better when the sample size is increasing
1. However, for delete-d, it does not show the similar pattern, the relative ratio becomew worse when n is large, which is what we expected. One factor could the number of covariates, that is when $p$ is large then it will be hard to make the jackknife work well??

#### GCTA with p = 100
```{r, echo = F}
data <- rbind(summary_final_GCTA_rr_8_d_1_p_100_n_50_1500,
              summary_final_GCTA_rr_8_d_d_p_100_n_50_1500,
              summary_final_GCTA_rr_8_d_.75_p_100_n_50_1500,
              summary_final_GCTA_rr_8_d_25_p_100_n_50_1500,
              fill = TRUE) %>% setorder(., by = d)

kable(data, "latex",  booktabs = T, linesep = "") %>%
kable_styling(latex_options = c("repeat_header", "scale_down")) 

data[,d := as.character(d)]
ggplot(data = data, 
       aes(x=n, y=relative_ratio, group=d)) +
  geom_line(aes(color=d)) +
  geom_point(aes(color=d)) +
  ggtitle("p = 100") + 
  ylim(c(0,3)) +
  theme_bw()
```

#### GCTA with p = 1000 
```{r, echo = F}
data <- rbind(summary_final_GCTA_rr_8_d_1_p_1000_n_50_1500,
              summary_final_GCTA_rr_8_d_d_p_1000_n_50_1500,
              summary_final_GCTA_rr_8_d_.75_p_1000_n_50_1500,
              summary_final_GCTA_rr_8_d_25_p_1000_n_50_1500,
              fill = TRUE) %>% setorder(., by = d)

kable(data, "latex",  booktabs = T, linesep = "") %>%
kable_styling(latex_options = c("repeat_header", "scale_down")) 

data[,d := as.character(d)]
ggplot(data = data, 
       aes(x=n, y=relative_ratio, group=d)) +
  geom_line(aes(color=d)) +
  geom_point(aes(color=d)) +
  ggtitle("p = 1000") + 
  ylim(c(0,3)) +
  theme_bw()
```

#### Eg with p = 100 
```{r, echo = F}
data <- rbind(summary_final_EigenPrism_8_d_1_p_100_n_50_1500,
              summary_final_EigenPrism_8_d_d_p_100_n_50_1500,
              summary_final_EigenPrism_8_d_.75_p_100_n_50_1500,
              summary_final_EigenPrism_8_d_25_p_100_n_50_1500,
              fill = TRUE) %>% setorder(., by = d)

kable(data, "latex",  booktabs = T, linesep = "") %>%
kable_styling(latex_options = c("repeat_header", "scale_down")) 
data[,d := as.character(d)]
ggplot(data = data, 
       aes(x=n, y=relative_ratio, group=d)) +
  geom_line(aes(color=d)) +
  geom_point(aes(color=d)) +
  ylim(c(0,3)) +
  ggtitle("p = 100") + 
  theme_bw()
```

#### Eg with p = 1000 
```{r, echo = F}
data <- rbind(summary_final_EigenPrism_8_d_1_p_1000_n_50_1500,
              summary_final_EigenPrism_8_d_d_p_1000_n_50_1500,
              summary_final_EigenPrism_8_d_.75_p_1000_n_50_1500,
              summary_final_EigenPrism_8_d_25_p_1000_n_50_1500,
              fill = TRUE) %>% setorder(., by = d)

kable(data, "latex",  booktabs = T, linesep = "") %>%
kable_styling(latex_options = c("repeat_header", "scale_down")) 
data[,d := as.character(d)]
ggplot(data = data, 
       aes(x=n, y=relative_ratio, group=d)) +
  geom_line(aes(color=d)) +
  geom_point(aes(color=d)) +
  ylim(c(0,3)) +
  ggtitle("p = 1000") + 
  theme_bw()
```

#### Dicker 2013 with p = 1000 
```{r, echo = F}
data <- rbind(summary_result_Dicker_jack_8_d_25_p_1000_n_50_1500,
              fill = TRUE) %>% setorder(., by = d)
kable(data, "latex",  booktabs = T, linesep = "") %>%
kable_styling(latex_options = c("repeat_header", "scale_down")) 
data[,d := as.character(d)]
ggplot(data = data, 
       aes(x=n, y=relative_ratio, group=d)) +
  geom_line(aes(color=d)) +
  geom_point(aes(color=d)) +
  ylim(c(0,3)) +
  ggtitle("p = 1000") + 
  theme_bw()
```