# Simulation study compare two GCTA and GCTA_rr
GCTA_rr is the `mixed.solve` function from `rrBLUP` r package.  
Based on the following simulation results, 

1. when $n<p$ case, those two methods' results are very closed to each other. 
1. when $n>p$ case, in terms of effect estimation and jackknife variance estimation those two methods's reuslts are similar to each other. But for the variance corrections are quite different. That is the statistics $Q$ of our method has a very large variance which leads to negative correction result.  

### setup 
- Independent 
- Normal
- $p = 100$
- $n = \{50, 75,100, 150, 200\}$
- with interaction terms
- main effect: $Var(X^T\beta) = \{0,8,100\}$ 
\newpage

```{r, include=F}
source("~/dev/projects/Chen_environmental_study/reports/proposed_GCTA_paper/est_var_analysis/est_combined_data/jackknife/GCTA_and_rr_v_jack_multiple_iteration.R")
```

### Simulation result 
### $Var(X^T\beta) = \{0\}$ 
```{r, }
summary_final_GCTA_0
summary_final_GCTA_rr_0
```

### $Var(X^T\beta) = \{100\}$ 
```{r, }
summary_final_GCTA_100
summary_final_GCTA_rr_100
```

### $Var(X^T\beta) = \{8\}$ 
```{r, }
summary_final_GCTA_8
summary_result_GCTA_500
summary_final_GCTA_rr_8
summary_result_GCTA_rr_500

```

### correlation test $ 
```{r, }
summary_final_cor_500
```

## compare the performance of delete 1 and delete d in variance estimation 
The delete-d jackknife varinace estimator is 
\[
\mathcal{v}_{J(d)} = \frac{n-d}{d} \cdot \frac{1}{S}\sum_{S}(\hat{\theta}_s - \hat{\theta}_{s.} )
\],
where $S =\binom{n}{d}$. 
Note that S could a very large value, so in the following simulation, only $S = 1000$ is used. In Jun Shao's another paper, he proposed an approximation of the deletel-d variance estimation. That is just select m from $S =\binom{n}{d}$ sub-samples and in that paper it recommended $m = n^{1.5}$.

### setup 
- Independent 
- Normal
- $p = \{100, 1000\}$
- $n = \{50, 75,100, 150, 200, 500, 750,1000, 1500, 2000\}$
- $d = 0.5 \times n$
- $n_{repeat} = 1000$ for delete d jackknife 
- main effect: $Var(X^T\beta) = 8$ 


\newpage

```{r, include = F}
source("~/dev/projects/Chen_environmental_study/reports/proposed_GCTA_paper/est_var_analysis/est_combined_data/jackknife/delete_d_GCTA_and_rr_v_jack_multiple_iteration.R")
```

### GCTA with p = 100 
```{r, echo = F}
rbind(summary_final_GCTA_8_d_1, summary_final_GCTA_8_d_d, summary_final_GCTA_8_d_d_m, fill = TRUE) %>% 
  kable(., "latex", longtable = T, booktabs = T) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 7)
```

### GCTA with p = 1000 
```{r, echo = F}
rbind(summary_final_GCTA_8_d_1_1000, summary_final_GCTA_8_d_d_1000) %>% 
  kable(., "latex", longtable = T, booktabs = T) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 7)
```

### GCTA_rr_rr with p = 100 
```{r, echo = F}
rbind(summary_final_GCTA_rr_8_d_1, summary_final_GCTA_rr_8_d_d, summary_final_GCTA_rr_8_d_d_m, fill = TRUE) %>% 
  kable(., "latex", longtable = T, booktabs = T) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 7)
```

### GCTA_rr with p = 1000 
```{r, echo = F}
rbind(summary_final_GCTA_rr_8_d_1_1000, summary_final_GCTA_rr_8_d_d_1000) %>% 
  kable(., "latex", longtable = T, booktabs = T) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 7)
```

### cor with n = 200
```{r, echo = F}
options(scipen=1, digits=5)
rbind(summary_final_cor_8_d_1_100, summary_final_cor_8_d_d_100) %>% 
  kable(., "latex", longtable = T, booktabs = T) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 7)
```

### median with n = 200
```{r, echo = F}
options(scipen=1, digits=5)
rbind(summary_final_median_8_d_1_1000, summary_final_median_8_d_d_1000) %>% 
  kable(., "latex", longtable = T, booktabs = T) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 7)
```
